---
title: "dataRbeautiful: Part 4"
author: "by [Matthew J. Oldach](https://github.com/moldach/) - `r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    css: 
    number_sections: FALSE
    includes:
      before_body: header.html
      after_body: footer.html
---
> Welcome to the final part of the series where I recreate data visualizations in R from the book "Knowledge is Beautiful" by David McCandless. 

David McCandless is author of two bestselling infographics books and gave a [TED talk about data visualization](http://www.ted.com/talks/david_mccandless_the_beauty_of_data_visualization). I bought his second book ["Knowledge is Beautiful"](https://informationisbeautiful.net/2014/knowledge-is-beautiful/), in 2015, which contains 196 beautiful infographics.

If you haven't checked out the first part of the series yet, please do.

# Plane Crashes
***
The **Plane Crashes** [dataset](https://docs.google.com/spreadsheets/d/1RT25s5oCsHJrahZkL6KWxAa2ZM9kD2aRakN9Cil_vHQ/edit#gid=11) is used for a couple of visualization which I will recreate. 

The first is a stacked-barplot showing the cause of crashes for every plane crash from 1993 to January 2017 where the flight hasn't been military, medical or a private chartered flight.

```{r}
library(tidyr)

df <- read.csv("worst_plane.csv")

# Drop the year plane model entered service
mini_df <- df %>% 
  select(-year_service) %>% 
# Gather the wide dataframe into a tidy format
  gather(key = cause, value = proportion, -plane)
# Order by cause
mini_df$cause <- factor(mini_df$cause, levels = c("human_error","weather", "mechanical", "unknown", "criminal"), ordered = TRUE)

# Create vector of plane names according to year they entered service
names <- unique(mini_df$plane)
names <- as.vector(names)
# sort by factor
mini_df$plane <- factor(mini_df$plane, levels = names)

ggplot(mini_df, aes(x=plane, y=proportion, fill=cause)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  # Reverse the order of a categorical axis
  scale_x_discrete(limits = rev(levels(mini_df$plane))) +
  # Select manual colors that McCandless used
  scale_fill_manual(values = c("#8E5A7E", "#A3BEC7", "#E1BD81", "#E9E4E0", "#74756F"), labels = c("Human Error", "Weather", "Mechanical", "Unknown", "Criminal")) +
  labs(title = "Worst Planes", caption = "Source: bit.ly/KIB_PlaneCrashes") +
  scale_y_reverse() +
  theme(legend.position = "right",
      panel.background = element_blank(),
        plot.title = element_text(size = 13, 
                                  family = "Georgia", 
                                  face = "bold", lineheight = 1.2),
      plot.caption = element_text(size = 5,
                                    hjust = 0.99, family = "Georgia"),
      axis.text = element_text(family = "Georgia"), 
      # Get rid of the x axis text/title 
      axis.text.x=element_blank(),
      axis.title.x=element_blank(),
      # and y axis title
      axis.title.y=element_blank(),
      # and legend title
      legend.title = element_blank(),
      legend.text = element_text(family = "Georgia"),
      axis.ticks = element_blank())
```

The second visualization is an alluvial diagram for which we can use the `ggalluvial` package.

```{r}
library(alluvial)
library(ggalluvial)

crash <- read.csv("crashes_alluvial.csv")

# stratum = cause, alluvium = freq

ggplot(crash, aes(weight = freq,
                  axis1 = phase,
                  axis2 = cause,
                  axis3 = total_crashes)) +
  geom_alluvium(aes(fill = cause),
                width = 0, knot.pos = 0, reverse = FALSE) +
  guides(fill = FALSE) +
  geom_stratum(width = 1/8, reverse = FALSE) +
  geom_text(stat = "stratum", label.strata = TRUE, reverse = FALSE, size = 2.5) +
  scale_x_continuous(breaks = 1:3, labels = c("phase", "causes", "total crashes")) +
  coord_flip() +
  labs(title = "Crash Cause", caption = "Source: bit.ly/KIB_PlaneCrashes") +
  theme(panel.background = element_blank(),
        plot.title = element_text(size = 13, 
                                  family = "Georgia", 
                                  face = "bold",
                                  lineheight = 1.2,
                                  vjust = -3,
                                  hjust = 0.05), 
        plot.caption = element_text(size = 5,
                                    hjust = 0.99, family = "Georgia"),  
        axis.text = element_text(family = "Georgia"),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.ticks.y = element_blank())

## Create the second Sankey Diagram

# Reorder dataframe
crash$total_crashes <- rep("YeS", 25)


library(magrittr)
library(magrittr)
# devtools::installgithub("Displayr/flipPlots")
library(flipPlots)


# Rearrange columns
crash %<>% select(total_crashes, cause, phase, freq)

SankeyDiagram(crash[, -4], link.color = "Source", label.show.varname = FALSE, weights = test_sankey$freq)
```

```{r}
gendergap <- read.csv("gendergap.csv")

# gather the dataset
tidy_gap <- gendergap %>% 
  gather(key = sex, value = salary, -title, -category)

category_means <- tidy_gap %>% 
  group_by(category) %>%
  summarize_at(vars(salary), mean)

category_width <- tidy_gap %>% 
  group_by(category) %>% 
  summarise(count = length(unique(title)))

q <- tidy_gap %>% ggplot(aes(x = title, y = salary, color = sex)) +
  facet_grid(. ~ category, scales = "free_x", space = "free") +
  geom_line(color = "white") +
  geom_point() +
  geom_blank() +
  scale_color_manual(values = c("#F49171", "#81C19C")) +
  geom_hline(data = category_means, aes(yintercept = salary), color = "white", alpha = 0.6, size = 1) +
  theme(legend.position = "none",
      panel.background = element_rect(color = "#242B47", fill = "#242B47"),
      plot.background = element_rect(color = "#242B47", fill = "#242B47"),
      axis.line = element_line(color = "grey48", size = 0.05, linetype = "dotted"),
      axis.text = element_text(family = "Georgia", color = "white"),
      axis.text.x = element_text(angle = 90),
      # Get rid of the y- and x-axis titles
      axis.title.y=element_blank(),
      axis.title.x=element_blank(),
      panel.grid.major.y = element_line(color = "grey48", size = 0.05),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_blank(),
      strip.background = element_rect(color = "#242B47", fill = "#242B47"),
      strip.text = element_text(color = "white", family = "Georgia"))

range_act <- range(range(tidy_gap$title))


```
